---
title: "function"
output: github_document
---
```{r}
library(faraway)
library(tidyverse)
library(dplyr)
library(broom)
library(glmnet)
library(olsrr)
library(MASS)
library(caret)
```
This will be model building procedure...
```{r}
sq_df = read_csv('squirrel_tidy.csv') %>% 
  janitor::clean_names()
```

```{r}
# longitudinal data function
long_lat_model =
  lm(cbind(long, lat) ~ shift + hectare_squirrel_number + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
summary(long_lat_model)
print(long_lat_model)

long_model =
  lm(long ~ hectare_squirrel_number + primary_fur_color + reaction + sounds, data = sq_df)
summary(long_model)
print(long_model)

long_model1 = 
  lm(cbind(long, lat) ~ hectare_squirrel_number + primary_fur_color + reaction + sounds, data = sq_df)
summary(long_model1)
print(long_model1)

```


# longitudinal model selection
model_criterion = lm(cbind(long, lat) ~ shift + hectare_squirrel_number + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
ols_step_best_subset(model_criterion)

model_stepwise <- lm(long ~ hectare_squirrel_number + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
ols_step_both_p(model_stepwise)



```{r}
# latitudinal data function
lat_model =
  lm(lat ~ hectare_squirrel_number + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
summary(lat_model)
```

```{r}
# latitudinal model selection
```

