---
title: "Central Park Data Analysis"
output: github_document
---
```{r, include = FALSE}
library(faraway)
library(tidyverse)
library(dplyr)
library(broom)
library(glmnet)
library(olsrr)
library(MASS)
library(caret)
library(modelr)
library(performance)
library(see)
library(patchwork)

options(digits = 7)
```

# Model Building
This will be model building procedure...

```{r, message = FALSE}
sq_df = read_csv('squirrel_tidy.csv') %>% 
  janitor::clean_names()
```

## P-value
The first step is to throw all the numerical variables into the model and check the p-value, the variables are shift + age + primary_fur_color + location + activity + reaction + sounds. Although hectare is also a numerical variable, it’s not included because the users of the model would not have the information of how many squirrels are there within a specific hectare, but they only have the information about the characteristics of specific squirrels that they want to look for. The variables with p-value less than 0.05 were removed from the model, and the model built with remaining variables was checked again to make sure that all of them had p-value less than 0.05. So, the first model candidate was produced with predictors being ‘shift’, ’ age’, ‘activity’, ‘reaction’, ‘sounds’.

```{r}
# longitudinal/latitudinal function
# raw data input (no hectare information because the user would not know the info. about it)
long_model =
  lm(long ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
summary(long_model)
print(long_model)

lat_model =
  lm(lat ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
summary(lat_model)
print(lat_model)
```


```{r}
# model selection using p-value
long_model1 = 
  lm(long ~ shift + age + activity + reaction + sounds, data = sq_df)
summary(long_model1)
print(long_model1)

lat_model1 =
  lm(lat ~ primary_fur_color + reaction + sounds, data = sq_df)
summary(lat_model1)
print(lat_model1)
```

## Automatic(Stepwise)
Then, we selected model using automatic procedure, specifically step-wise regression procedure. Backward, Forward or step-wise methods might produce different results, but we chose to use step-wise since it gives a single ‘best’ model. As the result, except for the location, all other 6 variables are included in this model, which is the second model candidate.

```{r}
# model selection using automatic procedure (forward/backward)
model_stepwise_long <- lm(long ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
ols_step_both_p(model_stepwise_long)

model_stepwise_lat <- lm(lat ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
ols_step_both_p(model_stepwise_lat)
```

## Criterion-based
Next, we used criterion-based procedure. The model with the largest adjusted R-square valued along with smallest AIC and BIC values are chosen to be the model candidate. It turned out that it also had all those 6 variables as the one in automatic procedure.

```{r}
# model selection using criterion-based procedure
model_criterion_long = lm(long ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
ols_step_best_subset(model_criterion_long)

model_criterion_lat <- lm(lat ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
ols_step_best_subset(model_criterion_lat)
```
## LASSO
LASSO model selection method was then used. After looking for the best lamda value, the third model candidate has all seven predictors, which means no variable was deleted from the selection procedure.

```{r}
# model selection using LASSO procedure: longitude
y_long <- sq_df$long
# hist(y_long)
x <- sq_df %>% 
  dplyr::select(
    shift, age, primary_fur_color,location, activity, reaction, sounds
    ) %>%
  data.matrix()

set.seed(2022)
cv_model_long <- cv.glmnet(x,y_long, alpha = 1)
best_lambda_long <- cv_model_long$lambda.min
long_lasso <- glmnet(x, y_long, alpha = 1, lambda = best_lambda_long)
coef(long_lasso)
```

```{r}
# model selection using LASSO procedure: latitude
y_lat <- sq_df$lat
# hist(y_lat)
x <- sq_df %>% 
  dplyr::select(
    shift, age, primary_fur_color,location, activity, reaction, sounds
    ) %>%
  data.matrix()

set.seed(2023)
cv_model_lat <- cv.glmnet(x,y_lat, alpha = 1)
best_lambda_lat <- cv_model_lat$lambda.min
lat_lasso <- glmnet(x, y_lat, alpha = 1, lambda = best_lambda_lat)
coef(lat_lasso)

```

# Model Selection
We have four different models as the final ‘best’ model candidate for now, and they are all nested within each other. We choose the ‘best’ model according to 3 criteria, the adjusted R-squared value, the cross-validation outcome(i.e. RMSE) and the rule of parsimony.


## Prediction models
The four different longitude-prediction model are as follows:

pvalue_fit_long = lm(long ~ shift + age + activity + reaction + sounds, data = sq_df)

auto_fit_long = lm(long ~ shift + hectare_squirrel_number + primary_fur_color + location + activity + reaction + sounds, data = sq_df)

crit_fit_long <- lm(long ~ hectare_squirrel_number + primary_fur_color + location + activity + reaction + sounds, data = sq_df)

lasso_fit_long = lm(long ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)

```{r}
# all longitudinal models
pvalue_fit_long = lm(long ~ shift + age + activity + reaction + sounds, data = sq_df)
pvalue_adjrsquared_long <- summary(pvalue_fit_long)$adj.r.squared

auto_fit_long = lm(long ~ sounds + primary_fur_color + reaction + activity + shift ,data = sq_df)
auto_adjrsquared_long <- summary(auto_fit_long)$adj.r.squared

crit_fit_long = lm(long ~ shift + age + primary_fur_color + activity + reaction + sounds, data = sq_df)
crit_adjrsquared_long <- summary(crit_fit_long)$adj.r.squared

lasso_fit_long = lm(long ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
lasso_adjrsquared_long <- summary(lasso_fit_long)$adj.r.squared
```
The four different latitude-prediction model are as follows:

pvalue_fit_lat = lm(lat ~ primary_fur_color + reaction + sounds, data = sq_df)

auto_fit_lat = lm(lat ~ sounds + primary_fur_color + reaction + activity + shift ,data = sq_df)

crit_fit_lat = lm(lat ~ shift + age + primary_fur_color + activity +  reaction + sounds, data = sq_df)

lasso_fit_lat = lm(lat ~ shift + age + primary_fur_color + activity + reaction + sounds, data = sq_df)

```{r}
# all latitudinal models
pvalue_fit_lat = lm(lat ~ primary_fur_color + reaction + sounds, data = sq_df)
pvalue_adjrsquared_lat <- summary(pvalue_fit_lat)$adj.r.squared

auto_fit_lat = lm(lat ~ sounds + primary_fur_color + reaction + activity + shift ,data = sq_df)
auto_adjrsquared_lat <- summary(auto_fit_lat)$adj.r.squared

crit_fit_lat = lm(lat ~ shift + age + primary_fur_color + activity +  reaction + sounds, data = sq_df)
crit_adjrsquared_lat <- summary(crit_fit_lat)$adj.r.squared

lasso_fit_lat = lm(lat ~ shift + age + primary_fur_color + activity + reaction + sounds, data = sq_df)
lasso_adjrsquared_lat <- summary(lasso_fit_lat)$adj.r.squared

```
## Cross Validation

The cross-validation divides model into the training set and the testing set. Models' comparison could be further done by comparing the cross-validation output: RMSE.

```{r}
# training and testing sets
cv_df <- crossv_mc(sq_df, 100)
cv_df <- cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

```{r}
# longitudinal model selection
cv_df_long = 
  cv_df %>% 
  mutate(
    pvalue_fit = map(train, ~lm(long ~ shift + age + activity + reaction + sounds, data = .x)),
    auto_fit = map(train, ~lm(long ~ sounds + primary_fur_color + reaction + activity + shift ,data = .x)),
    crit_fit = map(train, ~lm(long ~ shift + age + primary_fur_color + activity + reaction + sounds, data = .x)),
    lasso_fit = map(train, ~lm(long ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = .x))
  ) %>% 
  mutate(
    rmse_pvalue = map2_dbl(pvalue_fit, test, ~modelr::rmse(model = .x, data = .y)),
    rmse_auto = map2_dbl(auto_fit, test, ~modelr::rmse(model = .x, data = .y)),
    rmse_crit = map2_dbl(crit_fit, test, ~modelr::rmse(model = .x, data = .y)),
    rmse_lasso = map2_dbl(lasso_fit, test, ~modelr::rmse(model = .x, data = .y))
  )

# Comparing the RMSE plot

rplot_long <- cv_df_long %>% 
  dplyr::select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  labs(
    title = 'RMSE for Cross Validation of Longitudinal Models',
       x = 'models', y = 'RMSE'
  )+
  theme_bw()

# latitudinal model selection
cv_df_lat = 
  cv_df %>% 
  mutate(
    pvalue_fit = map(train, ~lm(lat ~ primary_fur_color + reaction + sounds, data = .x)),
    auto_fit = map(train, ~lm(lat ~ sounds + primary_fur_color + reaction + activity + shift ,data = .x)),
    crit_fit = map(train, ~lm(lat ~ shift + age + primary_fur_color + activity +  reaction + sounds, data = .x)),
    lasso_fit = map(train, ~lm(lat ~ shift + age + primary_fur_color + activity + reaction + sounds, data = .x))
  ) %>%
  mutate(
    rmse_pvalue = map2_dbl(pvalue_fit, test, ~modelr::rmse(model = .x, data = .y)),
    rmse_auto = map2_dbl(auto_fit, test, ~modelr::rmse(model = .x, data = .y)),
    rmse_crit = map2_dbl(crit_fit, test, ~modelr::rmse(model = .x, data = .y)),
    rmse_lasso = map2_dbl(lasso_fit, test, ~modelr::rmse(model = .x, data = .y))
  )

# Comparing the RMSE plot

rplot_lat <- cv_df_lat %>% 
  dplyr::select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  labs(
    title = 'RMSE for Cross Validation of Latitudinal Models',
       x = 'models', y = 'RMSE'
  )+
  theme_bw()

rplot_long + rplot_lat
```

## Visualization Tables
Visualization Tables are made for better comparison.
```{r}
# RMSE and adj Rsquared comparison table
methods <- c('pvalue','automatic','criterion-based','LASSO')
#longitude: parsimony since the RMSE and Rsquared are similar
RMSE_long <- 
  cv_df_long %>% 
  dplyr::select(starts_with('rmse_')) %>% 
  map(median) %>% 
  unlist()

adjrsquared_long <- c(pvalue_adjrsquared_long,auto_adjrsquared_long,
                       crit_adjrsquared_long, lasso_adjrsquared_long)
long_comparison <- tibble(
  methods,
  RMSE_long, 
  adjrsquared_long
)
long_comparison %>% knitr::kable()

# latitude: parsimony since the RMSE and Rsquared are similar
RMSE_lat <- 
  cv_df_lat %>% 
  dplyr::select(starts_with('rmse_')) %>% 
  map(median) %>% 
  unlist()

adjrsquared_lat <- c(pvalue_adjrsquared_lat,auto_adjrsquared_lat,
                       crit_adjrsquared_lat, lasso_adjrsquared_lat)
lat_comparison <- tibble(
  methods,
  RMSE_lat, 
  adjrsquared_lat
)
lat_comparison %>% knitr::kable()

```
# Final Models
## Final Longitudinal Model
The function of longitudinal has 5 predictors (shift + age + activity + reaction + sounds). Since the RMSE values and adjusted R-squared are approximately same among all functions, the principal of parsimony tells us to choose the the most succinct model. 
The formula of the latitudinal function is: longitudinal = -73.97 + 0.0005639 shift - 0.0009412 age - 0.0002267 activity + 0.0004809 reaction + 0.002142 sounds.

```{r}
# final model
long_model_final <- 
  lm(long ~ shift + age + activity + reaction + sounds, data = sq_df)
print(long_model_final)
long_model_final %>% broom::tidy()
```

The diagnosis plots of the final longitudinal prediction model:
```{r}
check_model(long_model_final, check = c("linearity", "outliers", "qq", "normality"))
```

## Final Latitudinal Model
The function of latitudinal has 3 predictors (primary_fur_color + reaction + sounds). Since the RMSE values and adjusted R-squared are approximately same among all functions, the principal of parsimony tells us to choose the the most succinct model. 
The formula of the latitudinal function is: longitudinal = 40.7814005- 0.0013140 primary_fur_color + 0.0007784 reaction + 0.0029072 sounds.

```{r}
lat_model_final <- 
  lm(lat ~ primary_fur_color + reaction + sounds, data = sq_df)
print(lat_model_final)
lat_model_final %>% broom::tidy()
```

The diagnosis plots of the final latitudinal prediction model:
```{r}
check_model(lat_model_final, check = c("linearity", "outliers", "qq", "normality"))
```
