---
title: "function"
output: github_document
---
```{r}
library(faraway)
library(tidyverse)
library(dplyr)
library(broom)
#library(glmnet)
library(olsrr)
library(MASS)
library(caret)
```
This will be model building procedure...
```{r}
sq_df = read_csv('squirrel_tidy.csv') %>% 
  janitor::clean_names()
```

```{r}
# longitudinal/latitudinal function
# raw data input (no hectare information because the user would not know the info. about it)
long_model =
  lm(long ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
summary(long_model)
print(long_model)

lat_model =
  lm(lat ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
summary(lat_model)
print(lat_model)
```

```{r}
# model selection using p-value
long_model1 = 
  lm(long ~ shift + age + activity + reaction + sounds, data = sq_df)
summary(long_model1)
print(long_model1)

lat_model1 =
  lm(lat ~ primary_fur_color + reaction + sounds, data = sq_df)
summary(lat_model1)
print(lat_model1)
```

```{r}
# model selection using automatic procedure (forward/backward)
model_stepwise_long <- lm(long ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
ols_step_both_p(model_stepwise_long)

model_stepwise_lat <- lm(lat ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
ols_step_both_p(model_stepwise_lat)
```

```{r}
# model selection using criterion-based procedure
model_criterion_long = lm(long ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
ols_step_best_subset(model_criterion_long)

model_criterion_lat <- lm(lat ~ shift + age + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
ols_step_best_subset(model_criterion_lat)
```

```{r}
# model selection using LASSO procedure
```

```{r}
set.seed(1)
# Use 10-fold validation 
train = trainControl(method = "cv", number = 10)
y1 <- data.frame(cbind(sq_df$long, sq_df$lat))
model_caret1 = train(long ~ primary_fur_color + reaction + sounds, data = sq_df,
trControl = train,
method = 'lm',
na.action = na.pass)

# fitting a regression line
model_caret1$finalModel
model_caret1
model_caret2 = train(lat ~ primary_fur_color + reaction + sounds, data = sq_df,
trControl = train,
method = 'lm',
na.action = na.pass)

# fitting a regression line
model_caret2$finalModel
model_caret2
```
# longitudinal model selection
model_criterion = lm(cbind(long, lat) ~ shift + hectare_squirrel_number + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
ols_step_best_subset(model_criterion)

model_stepwise <- lm(long ~ hectare_squirrel_number + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
ols_step_both_p(model_stepwise)



```{r}
# latitudinal data function
lat_model =
  lm(lat ~ hectare_squirrel_number + primary_fur_color + location + activity + reaction + sounds, data = sq_df)
summary(lat_model)
```

```{r}
# latitudinal model selection
```
