---
title: "Frequency"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(shiny)
library(patchwork)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE, 
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


```{r, include=FALSE}
census =
  read_csv("2018_Central_Park_Squirrel_Census_-_Squirrel_Data.csv")%>% 
  janitor::clean_names() %>%
  #filter(!is.na(primary_fur_color))%>%
  mutate(date=as.character(date),
         long = as.numeric(x),
         lat = as.numeric(y),
         shift = as.character(shift),
         activity = case_when( running == "TRUE" ~ "running",
                               chasing == "TRUE" ~ "chasing",
                               climbing == "TRUE" ~ "climbing",
                               eating == "TRUE" ~ "eating",
                               foraging == "TRUE" ~ "foraging"),
         reaction = case_when( approaches == "TRUE" ~ "approaches",
                               indifferent == "TRUE" ~ "indifferent",
                               runs_from == "TRUE" ~ "runs_from"),
         sounds = case_when( kuks == "TRUE" ~ "kuks",
                               quaas == "TRUE" ~ "quaas",
                               moans == "TRUE" ~ "moans"))%>%
  select(-x,-y,-running,-chasing,-climbing,-eating,-foraging,-above_ground_sighter_measurement,-color_notes,-other_activities,-specific_location ,-tail_flags ,-tail_twitches,-other_interactions,-kuks,-quaas,-moans,-approaches,-indifferent,-runs_from)
```

### Central Park Squirrels Distribution by Days (AM/PM)

In this part, we were interested in looking at the Central Park Squirrel Census collected in October 2018. We make this graph to analysis the frequency of Squirrel recorded for each of the observation (AM or PM).

```{r}
dates = 
  census %>%
  mutate(Month = substr(date, 1,2), Day = substr(date, 3,4), Year = substr(date,5,9))

dates %>% 
  group_by(Day, shift) %>%
  count() %>%
  ggplot() +
  geom_col(aes(x = Day, y = n, fill = shift)) +
  scale_fill_brewer(palette = "Paired") +
  labs(x = "Day", y = "Number of Observations", fill = 'Shift') +
  ggtitle('Central Park Squirrels Distribution by Days (AM/PM)') +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(fill='Time of day') 
```

### Central Park Squirrels Distribution by Primary Fur Colors

In this part, we were interested in the frequency of Squirrel recorded for each of the observation

```{r}
dates %>%
  group_by(Day, primary_fur_color) %>% 
  count() %>%
  ggplot() +
  geom_col(aes(x = Day, y = n, fill = primary_fur_color)) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle('Central Park Squirrels Distribution by Primary Fur Color') +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs( x='Day', y= 'Number of Observations') +
  labs(fill='Primary Fur Color') +
  scale_fill_manual(values = c("#000000", "#D2691E", "#D3D3D3", "white"))
```

### Central Park Squirrels Distribution by Age Group

```{r}
plot_1 = 
dates %>%
  filter(age !='?') %>%
  mutate(n=1) %>%
  group_by(age) %>%
  ggplot() +
  geom_col(aes(x = age, y = n, fill = shift), position="fill") +
  ggtitle('Demographics of Central Park Squirrels') +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs( x='Age Group', y= 'Proportion') +
  labs(fill='Time of day') +
  scale_fill_manual(values = c("#000000", "#D2691E", "#D3D3D3", "white"))
```

```{r}
plot_2 = 
dates %>%
  filter(age !='?') %>%
  mutate(n=1) %>%
  group_by(age) %>%
  ggplot() +
  geom_col(aes(y=n,x=age, fill=primary_fur_color), position="fill") +
  ggtitle('Demographics of Central Park Squirrels') +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs( x='Age Group', y= 'Proportion') +
  labs(fill='Primary Fur Color') +
  scale_fill_manual(values = c("#000000", "#D2691E", "#D3D3D3", "white"))
```

```{r}
plot_1 + plot_2
```

