---
title: "P8105 Project Proposal"
date: "November 11, 2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Team member!!
```{r, echo=FALSE, message=FALSE, warning=FALSE}
tibble(Name = c("Yimin Chen", "Feng Yan", "Jiong Ma", "Wenjing Yang", "Yang Yi"),
       UNI = c("yc4195", 
                "fy2297", 
                "jm5509",
                "wy2369",
                "yy3307"),)%>%
         knitr::kable()
```


### Tentative Title: Central Park Squirrel Data Analysis

### Data sources: 

- [City of New York Squirrel Census   2018](https://data.cityofnewyork.us/Environment/2018-Central-Park-Squirrel-Census-Squirrel-Data/vfnx-vebw) 



### Motivation
- Squirrels are everywhere, but is there a trend of where they prefer to stay with respect to their color, age, activities and all other features? Doing an analysis using squirrel census data may answer the question. Based on the results of the investigation we can compile relevant charts and make reasonable exploits, prevention or prediction based on information about squirrel habitat and activity patterns in Central Park.




### Intended final products

- A written report summarizing results of investigations
- A webpage containing motivation, introduction, questions and analyses, visualizations with interpretations and explanations (tables, graphs, interactive map, etc.), discussion, conclusions and next steps, limitations, and references. 

- Reasonable usage scenarios for the website can be provided, for instance:
Sometimes we might want to have a picnic or an event in Central Park but don't want to be disturbed by a ton of squirrels. Based on our results, a location with less squirrels could be discovered as the venue.
Researchers who want to collect data related to squirrel habits can accurately find the location of squirrels and related data through our results. Or he can directly extract and reference useful information about correlation of relationships between specific behaviors and age or color from our results.
- A video introducing our webpage in details




### Analyses, visualizations, and challenges

- analyses: 
Find location distribution of squirrels in Central Park in 2018.
Analyze the frequency for squirrel sightings by different fur colors.
Investigate the relationship of behavior observation and age & fur of squirrels.
ANOVA Test/ T-test- whether behavior are associated with age & fur of squirrels
Construct data model to predict the behavior/habitat of a specific squirrel 

- visualizations: 
Interactive map: Relative habitat of the squirrels in Central Park that match the filter
Scatter Plot (Smooth): Whether specific behaviors of squirrels generate (increase) trends according to age
Plot: Does the combination of squirrel's fur color produce a certain regularity of habitat or cluster effect?

- coding challenges: some information needs to be further interpreted in the raw dataset. For example, in the column of “Above Ground Sighter Measurement”, some entries are ‘FALSE’ or blank. Also, in the column of “Hectare”, what different entries represent for and how do we visualize it?



### Planned timeline

```{r, echo=FALSE, message=FALSE, warning=FALSE}
tibble(Date = c("Nov 3", "Nov 10", "Nov 15-18", "Nov 19", "Nov 25",  "Dec 1", "Dec 3", "Dec 6", "Dec 15"),
       Task = c("Brainstorming", 
                "Finishing and submitting proposal", 
                "Discuss about project validity",
                "Start coding!",
                "Progress review meeting and troubleshooting",
                "Finishing coding part and starting website part ",
                "Finishing building website",
                "Making screencast",
                "Presentation"),
       Due = c("NA", 
                "Nov 13", 
                "Nov 18",
                "NA",
                "Dec 10",
                "Dec 10",
                "Dec 10",
                "Dec 10",
                "Dec 15")) %>%
  knitr::kable()
```


