---
title: "shinny base"
author: "yimin chen"
date: "2022-11-29"
output: html_document
runtime: shiny
---

- analyses: 
Find location distribution of squirrels in Central Park in 2018.
Analyze the frequency for squirrel sightings by different fur colors.
Investigate the relationship of behavior observation and age & fur of squirrels.
ANOVA Test/ T-test- whether behavior are associated with age & fur of squirrels
Construct data model to predict the behavior/habitat of a specific squirrel 

- visualizations: 
Interactive map: Relative habitat of the squirrels in Central Park that match the filter
Scatter Plot (Smooth): Whether specific behaviors of squirrels generate (increase) trends according to age
Plot: Does the combination of squirrel's fur color produce a certain regularity of habitat or cluster effect?

---
title: "tidy and analysis"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
always_allow_html: true
---

- analyses: 
Find location distribution of squirrels in Central Park in 2018.
Analyze the frequency for squirrel sightings by different fur colors.
Investigate the relationship of behavior observation and age & fur of squirrels.
ANOVA Test/ T-test- whether behavior are associated with age & fur of squirrels
Construct data model to predict the behavior/habitat of a specific squirrel 

- visualizations: 
Interactive map: Relative habitat of the squirrels in Central Park that match the filter
Scatter Plot (Smooth): Whether specific behaviors of squirrels generate (increase) trends according to age
Plot: Does the combination of squirrel's fur color produce a certain regularity of habitat or cluster effect?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(scales)
library(ggthemes)
library(report)
library(sf)
library(shiny)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE, 
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r,include=FALSE}
census1 = #这是数据处理过后的，即编号0，1，2，3的，原始版在下面
  read_csv("~/Desktop/2018_Central_Park_Squirrel_Census_-_Squirrel_Data copy.csv")%>% 
  janitor::clean_names() %>%
  #filter(!is.na(primary_fur_color))%>%
  mutate(date=as.character(date),
         long = as.numeric(x),
         lat = as.numeric(y),
         shift = as.character(shift),
         primary_fur_color=as.character(primary_fur_color),
         activity = case_when( running == "TRUE" ~ "running",
                               chasing == "TRUE" ~ "chasing",
                               climbing == "TRUE" ~ "climbing",
                               eating == "TRUE" ~ "eating",
                               foraging == "TRUE" ~ "foraging"),
         reaction = case_when( approaches == "TRUE" ~ "approaches",
                               indifferent == "TRUE" ~ "indifferent",
                               runs_from == "TRUE" ~ "runs_from"),
         sounds = case_when( kuks == "TRUE" ~ "kuks",
                               quaas == "TRUE" ~ "quaas",
                               moans == "TRUE" ~ "moans"))%>%
         mutate(primary_fur_color = case_when(primary_fur_color == "Gray" ~ "1"
          ,primary_fur_color == "Cinnamon" ~ "2",
          primary_fur_color == "Black" ~ "3",
          TRUE ~ "0"
         ))%>%
  mutate(shift = case_when(shift == "AM" ~ "1"
          ,shift == "PM" ~ "2",
          TRUE ~ "0"
         ))%>%
  mutate(age = case_when(age == "Adult" ~ "1"
          ,age == "Juvenile" ~ "2",
          TRUE ~ "0"
         ))%>%
   mutate(activity = case_when(activity == "running" ~ "1"
          ,activity == "eating" ~ "2",
          activity == "foraging" ~ "3",
          activity == "climbing" ~ "4",
          activity == "chasing" ~ "5",
          TRUE ~ "0"
         ))%>%
  mutate(reaction = case_when(reaction == "indifferent" ~ "1"
          ,reaction == "runs_from" ~ "2",
          reaction == "approaches" ~ "3",
          TRUE ~ "0"
         ))%>%
  mutate(sounds = case_when(sounds == "kuks"~ "1",
                               sounds =="quaas" ~ "2",
                               sounds == "moans" ~ "3",
          TRUE ~ "0"
         ))%>%
  select(-x,-y,-running,-chasing,-climbing,-eating,-foraging,-above_ground_sighter_measurement,-color_notes,-other_activities,-specific_location ,-tail_flags ,-tail_twitches,-other_interactions,-kuks,-quaas,-moans,-approaches,-indifferent,-runs_from)
#write.csv(census1,"~/Desktop/squirrel_tidy.csv", row.names = TRUE)
```

```{r,include=FALSE}
census = #这是未数据处理过后的，即原始版
  read_csv("~/Desktop/2018_Central_Park_Squirrel_Census_-_Squirrel_Data copy.csv")%>% 
  janitor::clean_names() %>%
  #filter(!is.na(primary_fur_color))%>%
  mutate(date=as.character(date),
         long = as.numeric(x),
         lat = as.numeric(y),
         shift = as.character(shift),
         primary_fur_color=as.character(primary_fur_color),
         activity = case_when( running == "TRUE" ~ "running",
                               chasing == "TRUE" ~ "chasing",
                               climbing == "TRUE" ~ "climbing",
                               eating == "TRUE" ~ "eating",
                               foraging == "TRUE" ~ "foraging"),
         reaction = case_when( approaches == "TRUE" ~ "approaches",
                               indifferent == "TRUE" ~ "indifferent",
                               runs_from == "TRUE" ~ "runs_from"),
         sounds = case_when( kuks == "TRUE" ~ "kuks",
                               quaas == "TRUE" ~ "quaas",
                               moans == "TRUE" ~ "moans"))%>%
  select(-x,-y,-running,-chasing,-climbing,-eating,-foraging,-above_ground_sighter_measurement,-color_notes,-other_activities,-specific_location ,-tail_flags ,-tail_twitches,-other_interactions,-kuks,-quaas,-moans,-approaches,-indifferent,-runs_from)
#write.csv(census,"~/Desktop/squirrel_original.csv", row.names = TRUE)
#na.omit()
```

```{r}
skimr::skim(census)
glimpse(census)
view(census)
```


```{r}
nyc_squirrels1=
  census%>%
  ggplot(aes(long, lat)) +
  geom_point() +
  theme_map()
nyc_squirrels1
nyc_squirrelmap <- leaflet(data=census1) %>%
  addTiles() %>%  
  addMarkers(lng=~long, lat=~lat, popup= ~hectare)
nyc_squirrelmap
#这个要在shiny的格式里看，我以防万一还是用了markdown
```

```{r}
omitted<-na.omit(census)
omitted%>%
  filter(omitted$age!="?")
plot(factor(location) ~ factor(age), data = omitted)
plot(factor(location) ~ factor(shift), data = omitted)
plot(factor(location) ~ factor(activity), data = omitted)
plot(factor(location) ~ factor(reaction), data = omitted)
plot(factor(long) ~ factor(activity), data = omitted)
```


```{r}
library(revgeo)
revgeo(longitude=-73.968285, latitude=40.785091, provider = 'photon', output='frame')
library(rgdal)
d <- data.frame(lon=census$long, lat=census$lat)
coordinates(d) <- c("lon", "lat")
```

```{r}
# Plot the results.
par(mfrow=c(1,3))
plot.default(census$long,census$lat, main="Raw data", cex.axis=.95)
plot(d, axes=TRUE, main="Original lat-lon", cex.axis=.95)

#sf <- st_as_sf(d, coords = c('census$long', 'census$lat'))
#ggplot(sf) + 
  #geom_sf( )+
  #geom_point(aes(census$long, census$lat),color="steelblue") +
  #theme_map()
```

```{r}

```
