---
title: "Central Park Squirrels Data Analysis"
output:
  html_document:
    theme: cosmo
    toc : true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(shiny)
library(scales)
library(patchwork)
library(plotly)


knitr::opts_chunk$set(
 echo = TRUE,
 warning = FALSE,
 message = FALSE, 
 fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
sqr_nyc <- 
  read_csv('./squirrel_data_in_nyc/squirrel-data.csv') %>% 
  janitor::clean_names() %>% 
  mutate(
    lat = squirrel_latitude_dd_dddddd,
    long = squirrel_longitude_dd_dddddd,
    lat_long = str_c('POINT (', long, ' ',lat,')'),
    activity = tolower(activities),
    unique_squirrel_id = squirrel_id
  ) %>% 
  mutate(
    primary_fur_color = case_when(
      primary_fur_color == "Gray" ~ "gray",
      primary_fur_color == "Cinnamon" ~ "cinnamon",
      primary_fur_color == "Black" ~ "black",
      TRUE ~ "others"
         )
    )%>%
  mutate(
    activity = map(str_split(activity, ','), 1) %>% unlist(),
    activity = map(str_split(activity, ' '), 1) %>% unlist()
      )%>% 
  mutate(
    activity = case_when(
      activity == "running" ~ "running",
      activity == "eating" ~ "eating",
      activity == "foraging" ~ "foraging",
      activity == "climbing" ~ "climbing",
      activity == "chasing" ~ "chasing",
          TRUE ~ 'others'
         )
    )%>%
  mutate(
    location = map(str_split(location, ','), 1) %>% unlist()
      ) %>% 
  mutate(
    location = case_when(
    location == "Ground Plane"~ "gound plane",
    location =="Above Ground" ~ "above ground",
    TRUE ~ 'others'
         )
    ) %>%
  dplyr::select(
    -c("squirrel_id" ,"color_notes", "activities",
       "other_notes_or_observations",  "squirrel_longitude_dd_dddddd",
          "squirrel_latitude_dd_dddddd")
    ) %>% 
  dplyr::select(unique_squirrel_id,everything()) 

# sqr_nyc %>% group_by(activity) %>%
#   summarize(n = n()) %>%
#   filter(
#     activity %in% c('eating','foraging','climbing', 'chasing','running')
#     ) #332/433

#write_csv(sqr_nyc, file = './squirrel_data_in_nyc/sqr_nyc_clean.csv')
```

## Squirrel Activity by Primary Fur Color

In our dataset, there are 5 different kinds of activities reported which are foraging, running, eating, climbing and chasing. We make this plot to analysis central park squirrel activity by their primary fur color. 

```{r}
(plot_1 = 
  sqr_nyc %>%
  filter(!is.na(activity)) %>% 
  filter(!is.na(primary_fur_color)) %>% 
  group_by(primary_fur_color) %>%
  count(activity, sort = TRUE)%>%
  ggplot(aes(x = reorder(activity, n), y = n)) + 
  geom_bar(aes(fill = primary_fur_color), stat = "identity") + 
  scale_fill_manual(values = c("#000000", "#D2691E", "#D3D3D3", "white")) +
  theme_classic() + 
  facet_wrap(~primary_fur_color, nrow = 1) +
  labs(title = "Squirrel Activity by Primary Fur Color", y = 'Number of Observations', x = 'Activity') +
  coord_flip())

```


## Squirrel Activity by Location of Squirrel

We make this plot to analysis central park squirrel activity by their location (above ground, ground plane). 

```{r}
sqr_nyc %>%
  filter(!is.na(activity)) %>% 
  filter(!is.na(location)) %>% 
  group_by(activity) %>%
  mutate(n=1) %>%
  ggplot() +
  geom_col(aes(y=n,x = activity, fill = location), position="fill") +
  ggtitle('Central Park Squirrels Acitivities by Location') +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs( x = 'Activity', y= 'Proportion') +
  labs(fill = 'Location') 
```


## Primary Fur Color Map

```{r}
pal_coats <- colorFactor(c("#000000", "#D2691E", "#D3D3D3", "white"), domain = c("Black", "Cinnamon", "Grey", "NA"))

#Locations based on fur color
sqr_nyc %>%
  filter(!is.na(primary_fur_color)) %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(lng = ~long,
                   lat = ~lat, radius = 3, color = ~pal_coats(primary_fur_color), stroke = FALSE, fillOpacity = 0.5) %>%
  addLegend(position = "topright",pal = pal_coats, values = ~primary_fur_color)
```





