---
title: "tidy and analysis"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
always_allow_html: true
---

- analyses: 
Find location distribution of squirrels in Central Park in 2018.
Analyze the frequency for squirrel sightings by different fur colors.
Investigate the relationship of behavior observation and age & fur of squirrels.
ANOVA Test/ T-test- whether behavior are associated with age & fur of squirrels
Construct data model to predict the behavior/habitat of a specific squirrel 

- visualizations: 
Interactive map: Relative habitat of the squirrels in Central Park that match the filter
Scatter Plot (Smooth): Whether specific behaviors of squirrels generate (increase) trends according to age
Plot: Does the combination of squirrel's fur color produce a certain regularity of habitat or cluster effect?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(scales)
library(ggthemes)
library(report)
library(ggmap)
library(sf)
library(shiny)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE, 
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
census =
  read_csv("2018_Central_Park_Squirrel_Census_-_Squirrel_Data.csv")%>% 
  janitor::clean_names() %>%
  #filter(!is.na(primary_fur_color))%>%
  mutate(date=as.character(date),
         long = as.numeric(x),
         lat = as.numeric(y),
         shift = as.character(shift),
         activity = case_when( running == "TRUE" ~ "running",
                               chasing == "TRUE" ~ "chasing",
                               climbing == "TRUE" ~ "climbing",
                               eating == "TRUE" ~ "eating",
                               foraging == "TRUE" ~ "foraging"),
         reaction = case_when( approaches == "TRUE" ~ "approaches",
                               indifferent == "TRUE" ~ "indifferent",
                               runs_from == "TRUE" ~ "runs_from"),
         sounds = case_when( kuks == "TRUE" ~ "kuks",
                               quaas == "TRUE" ~ "quaas",
                               moans == "TRUE" ~ "moans"))%>%
  select(-x,-y,-running,-chasing,-climbing,-eating,-foraging,-above_ground_sighter_measurement,-color_notes,-other_activities,-specific_location ,-tail_flags ,-tail_twitches,-other_interactions,-kuks,-quaas,-moans,-approaches,-indifferent,-runs_from)
#na.omit()
skimr::skim(census)
glimpse(census)
view(census)
nyc_squirrels1=
  census%>%
  ggplot(aes(long, lat)) +
  geom_point() +
  theme_map()
nyc_squirrels1
nyc_squirrelmap <- leaflet(data=census) %>%
  addTiles() %>%  
  addMarkers(lng=~long, lat=~lat, popup= ~hectare)
nyc_squirrelmap
#这个要在shiny的格式里看，我以防万一还是用了markdown

omitted<-na.omit(census)
omitted%>%
  filter(omitted$age!="?")
plot(factor(location) ~ factor(age), data = omitted)
plot(factor(location) ~ factor(shift), data = omitted)
plot(factor(location) ~ factor(activity), data = omitted)
plot(factor(location) ~ factor(reaction), data = omitted)
plot(factor(long) ~ factor(activity), data = omitted)

library(revgeo)
revgeo(longitude=-73.968285, latitude=40.785091, provider = 'photon', output='frame')
library(rgdal)
d <- data.frame(lon=census$long, lat=census$lat)
coordinates(d) <- c("lon", "lat")

# Plot the results.
par(mfrow=c(1,3))
plot.default(census$long,census$lat, main="Raw data", cex.axis=.95)
plot(d, axes=TRUE, main="Original lat-lon", cex.axis=.95)


sf <- st_as_sf(d, coords = c('census$long', 'census$lat'))

```

```{r}

```
c('unique_squirrel_id', 'primary_fur_color', 'highlight_fur_color','combination_of_primary_and_highlight_color'))

ggplot(sf) + 
  geom_sf( )+
  geom_point(aes(census$long, census$lat),color="steelblue") +
  theme_map()